---
title: "Homework 11: Population Scale Analysis"
author: "Arshiya Zarmahd (PID: A16247996)"
format: pdf
---

## Section 4

> Q13. Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes.

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```

```{r}
nrow(expr)
```

462 individuals in this data file. 

```{r}
table(expr$geno)
```

This is the breakdown of how many individuals there are with each genotype.

```{r}
summary(expr)
```

Here is a quick summary of the data file. 

> Q14. Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3?

```{r}
library(ggplot2)
```

```{r}
ggplot(expr) + aes(geno, exp, fill=geno) + 
  geom_boxplot(notch=TRUE)
```

Here is a boxplot of the data dile. A|A is the highest expressed genotype, because it has the highest median, with G|G being the lowest expressed genotype. 





